Transforming Trash as an Artistic Act

### 2016-01-17 chapter1_final.tex
Print copy preperations, proof reading
Former-commit-id: c4e7e2d53515fdd32b70a48f44d38c9e790b69b9

\begin{singlespace}
\epigraph{One can even shout out through refuse \ldots}{\hfill---Kurt Schwitters, \textit{Kurt Schwitters}, 1985}
\end{singlespace}

@@

% Change in the production and consumption practices
\citet[11]{banash2013collage} draws attentions to the advancements in the recent history of humankind by stating that \quotes{over the course of the twentieth century, the twin developments of mass production and mass media in the capitalist economies of the [Western countries] completed a total transformation of everyday life, reorienting almost every activity toward consumption.} He argues that \quotes{things once locally produced and often handmade were now mass produced and commodified, turning local, artisanal producers into deskilled laborers serving the assembly line.} As a result of it amount of goods boost and the cost of them significantly reduced. Things once reused again and again, now thrown away because it is more affordable to replace with new one. As the production increased its by-product material waste is also increased. \quotes{The phenomenon of waste comes clearly into focus not merely as a by-product of manufacturing processes, but rather as an integral element in cycles of production and consumption} \citep[ix]{pye2010trashculture}.

@@

% Different scales
The scale of trash production is not same for everybody or every country. \citet[2]{chen2015waste} writes that by referring to analysis of World Bank \quotes{the amount of solid waste generated in cities is growing faster than the rate of urbanization.} She argues that when the people's income level and the rate of urbanization increased, production of waste also increases. Chen notes that nearly half of the world’s waste are produced by OECD\footnote{OECD (Organization for Economic Co-operation and Development) is an international economic organization whose members are mostly western developed countries.} countries. On the other hand, Africa and South Asia generate least waste. Similar to these as stated by \citet[16]{cerny1996recycled} \quotes{a person’s wealth has become measured not only in how much he or she can afford to consume, but in how much he or she can afford to throw away.} Moreover from trash many unexpected results can be extracted because as \citet[xxv]{zimring2012encyclopedia} argues that:

@@

% Different disciplines
Trash attracts the attention of different peoples and disciplines \quotes{ranging from economics to environmental studies, but most particularly by those studying consumerism or material culture} \citep[63]{emgin2012trashion}. For instance for archaeologist dumps near the residential area are one of the significant places to figure out what people consume in the ancient times \citep{rathje1992rubbish}. Moreover scholars argues that there is a relationship between technological and waste. Trash can be seen as a design and technological problem \citep{mcdonough2010cradle}. Further economical aspect of trash can not ignored. \citet[65]{emgin2012trashion} states that there is a huge business based on eco-friendly products and recyclable goods that is increasingly promoted.

% Different people
Various people including garbage collectors and artists interact with trash for different purposes. There are people who collects plastics, scraps and papers from waste bins and landfills. For the rag-pickers, collecting others excess and discard is significant source of economical income. They collect and classify recycle materials to sell them. By collecting and selling them they endure their life. On the other side for some of the people it is not considered as trash. As stated by the editor of the Garbage issue of ReVista, Christmas decorations at Chocó, a poor region on Colombia’s Pacific Coast, are \quotes{all crafted from used tin cans, old newspapers, discarded textiles and found wood objects} \citep{erlick2015editorsletter}. She realized that any of them called the practice as recycling. For them using trash again and again is very natural and it is part of their life. On the contrary for the developed countries trash considered as a thing that must be avoided. There is no place for trash in their life. It can be understood from this anecdote that approaches to the garbage are not same for all regions of the world. It can be said that \quotes{perceptions of waste and the value of material are neither static nor universally shared} \citep[xxvi]{zimring2012encyclopedia}.

@@

Thesis project focuses on disposable items and specifically paper trash. Its purpose is to transform trash and make it worth to reconsider. At this point I need to state that the process of transformation is not to be confused with the kind of industrial recycling that is applying automated procedures to the products such as paper cups, glass bottles, plastics caps and aluminum cans with the purpose of returning them to the industrial production as raw materials. Further many people are inside the business of transforming trash through recycling and reusing. People has been looking for aluminum cans, glass bottles and plastics on the streets. They collect and sell them to survive. There are companies that are collect waste materials to regain them to the industry. In thesis they are not in the scope. Artist that give attention to the trash and transform them is in the scope of this thesis.

% TODO relocation.
% MOVE TO Conclusion of art chapter, discussion of transformation.
% FROM Trashion: The Return of the Disposed by Bahar Emgin
%\ People are often eager to see objects that were once considered useless and tasteless when they have been invigorated with new life.

@@

%****************************************
\section{A Note on Terminology}
Many scholars and authors have used broad range of terms to refer the subject. Garbage, trash, rubbish, debris, detritus, waste, scrap, junk, refuse, discard, disposal, litter are some of them. Sometimes these words are used synonymously in causal speech but in fact there are slight differences between them. This situation signifies that the vocabulary of trash is so wide and there is no consensus on terminology. The authors of the book \quotes{Rubbish: the Archeology of Garbage} give clearer definition of some these words:

@@

Moreover artist Alice Bradshaw has published a 36 page \quotes{Rubbish Newspaper}. It is based on her extensive research on this subject. The newspaper contains these words and illustrates their place in literature and artworks.

>>>>

\begin{singlespace}
\epigraph{One can even shout out through refuse\ldots}{\hfill---Kurt Schwitters, \textit{Kurt Schwitters}, 1985}
\end{singlespace}

@@

% Change in the production and consumption practices
\citet[11]{banash2013collage} draws attentions to the advancements in the recent history of humankind by stating that \quotes{over the course of the twentieth century, the twin developments of mass production and mass media in the capitalist economies of the [western countries] completed a total transformation of everyday life, reorienting almost every activity toward consumption.} He argues that \quotes{things once locally produced and often handmade were now mass produced and commodified, turning local, artisanal producers into deskilled laborers serving the assembly line.} As a result of it amount of goods boost and the cost of them significantly reduced. Things once reused again and again, now thrown away because it is more affordable to replace with new one. As the production increased, its by-product material waste is also increased. \quotes{The phenomenon of waste comes clearly into focus not merely as a by-product of manufacturing processes, but rather as an integral element in cycles of production and consumption} \citep[ix]{pye2010trashculture}.

@@

% Different scales
The scale of trash production is not same for everybody or every country. \citet[2]{chen2015waste} writes that by referring to analysis of World Bank \quotes{the amount of solid waste generated in cities is growing faster than the rate of urbanization.} She argues that when the people’s income level and the rate of urbanization increased, production of waste also increases. Chen notes that nearly half of the world’s waste are produced by OECD\footnote{OECD (Organization for Economic Co-operation and Development) is an international economic organization whose members are mostly western developed countries.} countries. On the other hand, Africa and South Asia generate least waste. Similar to these as stated by \citet[16]{cerny1996recycled} \quotes{a person’s wealth has become measured not only in how much he or she can afford to consume, but in how much he or she can afford to throw away.} Moreover from trash many unexpected results can be extracted because as \citet[xxv]{zimring2012encyclopedia} argues that:

@@

% Different disciplines
Trash attracts the attention of different peoples and disciplines \quotes{ranging from economics to environmental studies, but most particularly by those studying consumerism or material culture} \citep[63]{emgin2012trashion}. For instance for archaeologist dumps near the residential area are one of the significant places to figure out what people consume in the ancient times \citep{rathje1992rubbish}. Moreover scholars argues that there is a relationship between technological and waste. Trash can be seen as a design and technological problem \citep{mcdonough2010cradle}. Further economic aspect of trash can not be ignored. \citet[65]{emgin2012trashion} states that there is a huge business based on eco-friendly products and recyclable goods that is increasingly promoted.

% Different people
Various people including garbage collectors and artists interact with trash for different purposes. There are people who collects plastics, scraps and papers from waste bins and landfills. For the rag-pickers, collecting others excess and discard is significant source of economic income. They collect and classify recycle materials to sell them. By collecting and selling them they endure their life. On the other side for some of the people it is not considered as trash. As stated by the editor of the Garbage issue of ReVista, Christmas decorations at Chocó, a poor region on Colombia’s Pacific Coast, are \quotes{all crafted from used tin cans, old newspapers, discarded textiles and found wood objects} \citep{erlick2015editorsletter}. She realized that any of them called the practice as recycling. For them using trash again and again is very natural and it is part of their life. On the contrary for the developed countries trash considered as a thing that must be avoided. There is no place for trash in their life. It can be understood from this anecdote that approaches to the garbage are not same for all regions of the world. It can be said that \quotes{perceptions of waste and the value of material are neither static nor universally shared} \citep[xxvi]{zimring2012encyclopedia}.

@@

Thesis project focuses on disposable items and specifically paper trash. Its purpose is to transform trash and make it worth to reconsider. At this point I need to state that the process of transformation is not to be confused with the kind of industrial recycling that is applying automated procedures to the products such as paper cups, glass bottles, plastics caps and aluminum cans with the purpose of returning them to the industrial production as raw materials. Further many people are inside the business of transforming trash through recycling and reusing. People has been looking for aluminum cans, glass bottles and plastics on the streets and garbage bins. They collect and sell them for economic purposes. There are companies that are collect waste materials to regain them to the industry. In thesis they are not in the scope. Artist that give attention to the trash and transform them is in the scope of this thesis.

@@

%****************************************
\section{A Note on Terminology}
Many scholars and authors have used broad range of terms to refer the subject. Garbage, trash, rubbish, debris, detritus, waste, scrap, junk, refuse, discard, disposal, litter are some of them. Sometimes these words are used synonymously in causal speech but in fact there are slight differences between them. This situation signifies that the vocabulary of trash is so wide and there is no consensus on terminology. The authors of the book \textit{Rubbish: the Archeology of Garbage} give clearer definition of some these words:

@@

Moreover artist Alice Bradshaw has published a 36 page \textit{Rubbish Newspaper}. It is based on her extensive research on this subject. The newspaper contains these words and illustrates their place in literature and artworks.
